---
output: 
  github_document:
    toc: false
    fig_width: 10.08
    fig_height: 6
tags: [r, ecg_preprocessing]
vignette: >
  %\VignetteIndexEntry{README}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
csl: utils/apa.csl
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# options and parameters
options(digits=3)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi=450,
  fig.path = "../../studies/ecg_benchmark/figures/"
)

# Setup python - you need to change the path to your python distribution
library(reticulate)
# reticulate::use_python("D:/Downloads/WPy64-3810/python-3.8.1.amd64/")
reticulate::use_python("C:/Program Files/Python39/python.exe")
matplotlib <- import("matplotlib")
matplotlib$use("Agg", force = TRUE)
```




# Benchmarking of ECG Preprocessing Methods

*This study can be referenced by* [*citing the package*](https://github.com/neuropsychology/NeuroKit#citation).

**We'd like to publish this study, but unfortunately we currently don't have the time. If you want to help to make it happen, please contact us!**

## Introduction

This work is a replication and extension of the work by @porr2019r, that compared the performance of different popular R-peak detectors.


## Databases

### Glasgow University Database

The GUDB Database [@howell2018high] contains ECGs from 25 subjects. Each subject was recorded performing 5 different tasks for two minutes (sitting, doing a maths test on a tablet, walking on a treadmill, running on a treadmill, using a hand bike). The sampling rate is 250Hz for all the conditions.

The script to download and format the database using the [**ECG-GUDB**](https://github.com/berndporr/ECG-GUDB) Python package by Bernd Porr can be found [**here**](https://github.com/neuropsychology/NeuroKit/blob/dev/data/gudb/download_gudb.py).

### MIT-BIH Arrhythmia Database

The MIT-BIH Arrhythmia Database [MIT-Arrhythmia; @moody2001impact] contains 48 excerpts of 30-min of two-channel ambulatory ECG recordings sampled at 360Hz and 25 additional recordings from the same participants including common but clinically significant arrhythmias (denoted as the `MIT-Arrhythmia-x` database).

The script to download and format the database using the can be found [**here**](https://github.com/neuropsychology/NeuroKit/blob/dev/data/mit_arrhythmia/download_mit_arrhythmia.py).

<!-- ### MIT-BIH Noise Stress Test Database -->

<!-- The MIT-BIH Noise Stress Test Database [MIT-NST; @moody1984bih] features two 30-minute recordings distorted by adding different types and levels of synthesized noise typical of electrode motion artefacts. -->


### MIT-BIH Normal Sinus Rhythm Database

This database includes 18 clean long-term ECG recordings of subjects. Due to memory limits, we only kept the second hour of recording of each participant.

The script to download and format the database using the can be found [**here**](https://github.com/neuropsychology/NeuroKit/blob/dev/data/mit_normal/download_mit_normal.py).


### Lobachevsky University Electrocardiography Database

The Lobachevsky University Electrocardiography Database [LUDB; @kalyakulina2018lu] consists of 200 10-second 12-lead ECG signal records representing different morphologies of the ECG signal. The ECGs were collected from healthy volunteers and patients, which had various cardiovascular diseases. The boundaries of P, T waves and QRS complexes were manually annotated by cardiologists for all 200 records.

### Fantasia Database

The Fantasia database [@iyengar1996age] consists of twenty young and twenty elderly healthy subjects. All subjects remained in a resting state in sinus rhythm while watching the movie Fantasia (Disney, 1940) to help maintain wakefulness. The continuous ECG signals were digitized at 250 Hz. Each heartbeat was annotated using an automated arrhythmia detection algorithm, and each beat annotation was verified by visual inspection.




### Concanate them together

```{python, eval=FALSE}
import pandas as pd

# Load ECGs
ecgs = ["../../data/gudb/ECGs.csv",
        "../../data/mit_arrhythmia/ECGs.csv",
        "../../data/mit_normal/ECGs.csv",
        "../../data/ludb/ECGs.csv",
        "../../data/fantasia/ECGs.csv"]

# Load True R-peaks location
rpeaks = [pd.read_csv("../../data/gudb/Rpeaks.csv"),
          pd.read_csv("../../data/mit_arrhythmia/Rpeaks.csv"),
          pd.read_csv("../../data/mit_normal/Rpeaks.csv"),
          pd.read_csv("../../data/ludb/Rpeaks.csv"),
          pd.read_csv("../../data/fantasia/Rpeaks.csv")]
```




## Study 1: Comparing Different R-Peaks Detection Algorithms

### Procedure

#### Setup Functions
```{python, eval=FALSE}
import neurokit2 as nk

def neurokit(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]

def pantompkins1985(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="pantompkins1985")
    return info["ECG_R_Peaks"]
    
def hamilton2002(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="hamilton2002")
    return info["ECG_R_Peaks"]
    
def martinez2003(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="martinez2003")
    return info["ECG_R_Peaks"]
    
def christov2004(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="christov2004")
    return info["ECG_R_Peaks"]
    
def gamboa2008(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="gamboa2008")
    return info["ECG_R_Peaks"]

def elgendi2010(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="elgendi2010")
    return info["ECG_R_Peaks"]
    
def engzeemod2012(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="engzeemod2012")
    return info["ECG_R_Peaks"]
    
def kalidas2017(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="kalidas2017")
    return info["ECG_R_Peaks"]
    
def rodrigues2020(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="rodrigues2020")
    return info["ECG_R_Peaks"]
```

#### Run the Benchmarking

*Note: This takes a long time (several hours).*

```{python, eval=FALSE}
results = []
for method in [neurokit, pantompkins1985, hamilton2002, martinez2003, christov2004, 
               gamboa2008, elgendi2010, engzeemod2012, kalidas2017, rodrigues2020]:
    for i in range(len(rpeaks)):
        data_ecg = pd.read_csv(ecgs[i])
        result = nk.benchmark_ecg_preprocessing(method, data_ecg, rpeaks[i])
        result["Method"] = method.__name__
        results.append(result)
results = pd.concat(results).reset_index(drop=True)

results.to_csv("data_detectors.csv", index=False)
```










### Results

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(easystats)
library(lme4)

data <- read.csv("data_detectors.csv", stringsAsFactors = FALSE) %>% 
  mutate(Method = fct_relevel(Method, "neurokit", "pantompkins1985", "hamilton2002", "martinez2003", "christov2004", "gamboa2008", "elgendi2010", "engzeemod2012", "kalidas2017", "rodrigues2020"))

colors <- c("neurokit"="#E91E63", "pantompkins1985"="#f44336", "hamilton2002"="#FF5722", "martinez2003"="#FF9800", "christov2004"="#FFC107", "gamboa2008"="#4CAF50", "elgendi2010"="#009688", "engzeemod2012"="#2196F3", "kalidas2017"="#3F51B5", "rodrigues2020"="#9C27B0") 
```

#### Errors and bugs

```{r, warning=FALSE, message=FALSE, fig.height=14, fig.width=8}
data %>% 
  mutate(Error = case_when(
    Error == "index -1 is out of bounds for axis 0 with size 0" ~ "index -1 out of bounds",
    Error == "index 0 is out of bounds for axis 0 with size 0" ~ "index 0 out of bounds",
    TRUE ~ Error)) %>% 
  group_by(Database, Method) %>% 
  mutate(n = n()) %>% 
  group_by(Database, Method, Error) %>% 
  summarise(Percentage = n() / unique(n)) %>% 
  ungroup() %>% 
  mutate(Error = fct_relevel(Error, "None")) %>% 
  ggplot(aes(x=Error, y=Percentage, fill=Method)) +
    geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
    facet_wrap(~Database, nrow=5) +
    theme_modern() +
    theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1)) +
    scale_fill_manual(values=colors)
```


**Conclusion:** It seems that `gamboa2008` and `martinez2003` are particularly prone to errors, especially in the case of a noisy ECG signal. Aside from that, the other algorithms are quite resistant and bug-free.


```{r, warning=FALSE, message=FALSE}
data <- filter(data, Error == "None")
data <- filter(data, !is.na(Score))
```

#### Computation Time

##### Descriptive Statistics

```{r, warning=FALSE, message=FALSE}
# Normalize duration
data <- data %>% 
  mutate(Duration = (Duration) / (Recording_Length * Sampling_Rate)) 

data %>% 
  ggplot(aes(x=Method, y=Duration, fill=Method)) +
    geom_jitter2(aes(color=Method, group=Database), size=3, alpha=0.2, position=position_jitterdodge()) +
    geom_boxplot(aes(alpha=Database), outlier.alpha = 0) +
    geom_hline(yintercept=0, linetype="dotted") +
    theme_modern() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_alpha_manual(values=seq(0, 1, length.out=8)) +
    scale_color_manual(values=colors) +
    scale_fill_manual(values=colors) +
    scale_y_sqrt() +
    ylab("Duration (seconds per data sample)")
```



<!-- ```{r, warning=FALSE, message=FALSE} -->
<!-- # Interaction time and recording length -->
<!-- data %>%  -->
<!--   ggplot(aes(x=Recording_Length, y=Duration, color=Method)) + -->
<!--     geom_jitter(aes(shape=Database), size=3, alpha=0.7) + -->
<!--     geom_smooth(method="lm", se = FALSE) + -->
<!--     theme_modern() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust = 1)) + -->
<!--     scale_color_manual(values=colors) + -->
<!--     scale_fill_manual(values=colors) + -->
<!--     scale_y_sqrt() -->
<!-- ``` -->



##### Statistical Modelling

```{r, warning=FALSE, message=FALSE}
model <- lmer(Duration ~ Method + (1|Database) + (1|Participant), data=data)

means <- modelbased::estimate_means(model)

arrange(means, Mean)

means %>% 
  ggplot(aes(x=Method, y=Mean, color=Method)) +
  geom_line(aes(group=1), size=1) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high), size=1) +
  geom_hline(yintercept=0, linetype="dotted") +
  theme_modern() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values=colors) +
    ylab("Duration (seconds per data sample)")
```


**Conclusion:** It seems that `gamboa2008` and `neurokit` are the fastest methods, followed by `martinez2003`, `kalidas2017`, `rodrigues2020` and `hamilton2002`. The other methods are then substantially slower.




#### Accuracy

**Note:** The accuracy is computed as the absolute distance from the original "true" R-peaks location. As such, the closest to zero, the better the accuracy.


##### Descriptive Statistics


```{r, warning=FALSE, message=FALSE}
data <- data %>% 
  mutate(Outlier = performance::check_outliers(Score, threshold = list(zscore = stats::qnorm(p = 1 - 0.000001)))) %>% 
  filter(Outlier == 0)

data %>% 
  ggplot(aes(x=Database, y=Score)) +
    geom_boxplot(aes(fill=Method), outlier.alpha = 0, alpha=1) +
    geom_jitter2(aes(color=Method, group=Method), size=3, alpha=0.2, position=position_jitterdodge()) +
    geom_hline(yintercept=0, linetype="dotted") +
    theme_modern() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_color_manual(values=colors) +
    scale_fill_manual(values=colors) +
    scale_y_sqrt() +
    ylab("Amount of Error") 
```


##### Statistical Modelling

```{r, warning=FALSE, message=FALSE}
model <- lmer(Score ~ Method + (1|Database) + (1|Participant), data=data)

means <- modelbased::estimate_means(model)

arrange(means, abs(Mean))

means %>% 
  ggplot(aes(x=Method, y=Mean, color=Method)) +
  geom_line(aes(group=1), size=1) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high), size=1) +
  geom_hline(yintercept=0, linetype="dotted") +
  theme_modern() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values=colors) +
    ylab("Amount of Error") 
```

**Conclusion:** It seems that `neurokit`, `kalidas2017` and `martinez2003` the most accurate algorithms to detect R-peaks. This pattern of results differs a bit from @porr2019r that outlines `engzeemod2012`, `elgendi2010`, `kalidas2017` as the most accurate and `christov2004`, `hamilton2002` and `pantompkins1985` as the worse. Discrepancies could be due to the differences in data and analysis, as here we used more databases and modelled them by respecting their hierarchical structure using mixed models.

### Conclusion

Based on the accuracy / execution time criterion, it seems like `neurokit` is the best R-peak detection method, followed by `kalidas2017`.









## Study 2: Normalization

### Procedure


#### Setup Functions
```{python, eval=FALSE}
import neurokit2 as nk

def none(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]

def mean_detrend(ecg, sampling_rate):
    ecg = nk.signal_detrend(ecg, order=0)
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]
    
def standardize(ecg, sampling_rate):
    ecg = nk.standardize(ecg)
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]
```

#### Run the Benchmarking

*Note: This takes a long time (several hours).*

```{python, eval=FALSE}
results = []
for method in [none, mean_detrend, standardize]:
    for i in range(len(rpeaks)):
        data_ecg = pd.read_csv(ecgs[i])
        result = nk.benchmark_ecg_preprocessing(method, data_ecg, rpeaks[i])
        result["Method"] = method.__name__
        results.append(result)
results = pd.concat(results).reset_index(drop=True)

results.to_csv("data_normalization.csv", index=False)
```


### Results

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(easystats)
library(lme4)

data <- read.csv("data_normalization.csv", stringsAsFactors = FALSE) %>% 
  mutate(Database = ifelse(str_detect(Database, "GUDB"), paste0(str_replace(Database, "GUDB_", "GUDB ("), ")"), Database),
         Method = fct_relevel(Method, "none", "mean_removal", "standardization"),
         Participant = paste0(Database, Participant)) %>% 
  filter(Error == "None") %>% 
  filter(!is.na(Score))

colors <- c("none"="#607D8B", "mean_removal"="#673AB7", "standardization"="#00BCD4") 
```

#### Accuracy


##### Descriptive Statistics


```{r, warning=FALSE, message=FALSE}
data <- data %>% 
  mutate(Outlier = performance::check_outliers(Score, threshold = list(zscore = stats::qnorm(p = 1 - 0.000001)))) %>% 
  filter(Outlier == 0)

data %>% 
  ggplot(aes(x=Database, y=Score)) +
    geom_boxplot(aes(fill=Method), outlier.alpha = 0, alpha=1) +
    geom_jitter2(aes(color=Method, group=Method), size=3, alpha=0.2, position=position_jitterdodge()) +
    geom_hline(yintercept=0, linetype="dotted") +
    theme_modern() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_color_manual(values=colors) +
    scale_fill_manual(values=colors) +
    scale_y_sqrt() +
    ylab("Amount of Error") 
```

##### Statistical Modelling

```{r, warning=FALSE, message=FALSE}
model <- lmer(Score ~ Method + (1|Database) + (1|Participant), data=data)

modelbased::estimate_contrasts(model) 

means <- modelbased::estimate_means(model)
arrange(means, abs(Mean))

means %>% 
  ggplot(aes(x=Method, y=Mean, color=Method)) +
  geom_line(aes(group=1), size=1) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high), size=1) +
  theme_modern() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values=colors) +
    ylab("Amount of Error") 
```

### Conclusion

No significant benefits added by normalization for the `neurokit` method.



## Study 3: Low Frequency Trends Removal


### Procedure

#### Setup Functions
```{python, eval=FALSE}
import neurokit2 as nk

def none(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]

# Detrending-based
def polylength(ecg, sampling_rate):
    length = len(ecg) / sampling_rate
    ecg = nk.signal_detrend(ecg, method="polynomial", order=int(length / 2))
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]
    
def tarvainen(ecg, sampling_rate):
    ecg = nk.signal_detrend(ecg, method="tarvainen2002")
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]
    
def locreg(ecg, sampling_rate):
    ecg = nk.signal_detrend(ecg, 
                            method="locreg", 
                            window=0.5*sampling_rate, 
                            stepsize=0.02*sampling_rate)
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]
    
def rollingz(ecg, sampling_rate):
    ecg = nk.standardize(ecg, window=sampling_rate*2)
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]
    
# Filtering-based
```

#### Run the Benchmarking

*Note: This takes a very long time (several hours).*

```{python, eval=FALSE}
results = []
for method in [none, polylength, tarvainen, locreg, rollingz]:
    result = nk.benchmark_ecg_preprocessing(method, ecgs, rpeaks)
    result["Method"] = method.__name__
    results.append(result)
results = pd.concat(results).reset_index(drop=True)

results.to_csv("data_lowfreq.csv", index=False)
```


## References
