---
output: 
  github_document:
    toc: false
    fig_width: 10.08
    fig_height: 6
tags: [r, ecg_preprocessing]
vignette: >
  %\VignetteIndexEntry{README}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# options and parameters
options(digits=3)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi=450,
  fig.path = "figures/"
)

# Setup python - you need to change the path to your python distribution
library(reticulate)
reticulate::use_python("D:/Downloads/WPy64-3810/python-3.8.1.amd64/")
matplotlib <- import("matplotlib")
matplotlib$use("Agg", force = TRUE)
```




# Benchmarking of ECG Preprocessing Methods

## Databases

### GUDB (Howell \& Porr, 2018)

The GUDB Database (Howell \& Porr, 2018) contains ECGs from 25 subjects. Each subject was recorded performing 5 different tasks for two minutes (sitting, doing a maths test on a tablet, walking on a treadmill, running on a treadmill, using a hand bike). The sampling rate is 250Hz for all the conditions.

The script to download and format the database using the [**ECG-GUDB**](https://github.com/berndporr/ECG-GUDB) Python package by Bernd Porr can be found [**here**]().

### WFDB 

**TODO.**

### Concanate them together

```{python}
import pandas as pd

# ECG
ecgs_gudb = pd.read_csv("../../data/gudb/ECGs.csv")
# ecgs_wfdb = NA

ecgs = ecgs_gudb

# True R-peaks location
rpeaks_gudb = pd.read_csv("../../data/gudb/Rpeaks.csv")
# rpeaks_wfdb = NA
rpeaks = rpeaks_gudb
```







## Comparing Different R-Peaks Detection Algorithms

### Setup Functions
```{python}
import neurokit2 as nk

def neurokit(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="neurokit")
    return info["ECG_R_Peaks"]

def gamboa2008(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="gamboa2008")
    return info["ECG_R_Peaks"]

def engzeemod2012(ecg, sampling_rate):
    signal, info = nk.ecg_peaks(ecg, sampling_rate=sampling_rate, method="engzeemod2012")
    return info["ECG_R_Peaks"]
```

### Run the Benchmark
```{python}
results = []
for method in [neurokit, gamboa2008, engzeemod2012]:
    result = nk.benchmark_ecg_preprocessing(method, r["ecgs"], r["rpeaks"])
    result["Method"] = method.__name__
    results.append(result)
results = pd.concat(results).reset_index(drop=True)
```










## Results

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(easystats)

data <- py$results %>% 
  mutate(Database = ifelse(str_detect(Database, "GUDB"), paste0(str_replace(Database, "GUDB_", "GUDB ("), ")"), Database))
colors <- c("neurokit"="#E91E63", "gamboa2008"="#2196F3", "engzeemod2012"="#FF9800")
```

### Errors

```{r, warning=FALSE, message=FALSE}
data %>% 
  group_by(Database, Method) %>% 
  mutate(n = n()) %>% 
  group_by(Database, Method, Error) %>% 
  summarise(Percentage = n() / unique(n)) %>% 
  ungroup() %>% 
  mutate(Error = fct_relevel(Error, "None"),
         Method = fct_relevel(Method, "neurokit")) %>% 
  ggplot(aes(x=Error, y=Percentage, fill=Method)) +
    geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
    facet_grid(~Database) +
    theme_modern() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values=colors)

data <- filter(data, Error == "None")
```

### Time

```{r, warning=FALSE, message=FALSE}
data %>% 
  ggplot(aes(x=Database, y=Duration)) +
    geom_boxplot(aes(fill=Method), outlier.alpha = 0, alpha=0.5) +
    geom_jitter2(aes(color=Method, group=Method), size=3, alpha=0.8, position=position_jitterdodge()) +
    theme_modern() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_color_manual(values=colors) +
    scale_fill_manual(values=colors)
```

### Score

```{r, warning=FALSE, message=FALSE}
data %>% 
  ggplot(aes(x=Database, y=Score)) +
    geom_boxplot(aes(fill=Method), outlier.alpha = 0, alpha=0.5) +
    geom_jitter2(aes(color=Method, group=Method), size=3, alpha=0.8, position=position_jitterdodge()) +
    theme_modern() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_color_manual(values=colors) +
    scale_fill_manual(values=colors) +
    scale_y_sqrt() +
    ylab("Amount of Error")
```


# References

Howell, L., & Porr, B. (2018). High precision ECG Database with annotated R peaks, recorded and filmed under realistic conditions.
