---
title: '**NeuroKit2: A Python Toolbox for Neurophysiological Signal Processing**'
shorttitle        : "NeuroKit2"
author:
  - name          : "Dominique Makowski"
    affiliation   : " 1,*"
    corresponding : no    # Define only one corresponding author
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
  - name          : "Tam Pham"
    affiliation   : " 1"
  - name          : "Zen J. Lau"
    affiliation   : " 1"
  - name          : "Jan C. Brammer"
    affiliation   : " 2"
  - name          : "Francois Lespinasse"
    affiliation   : " 3"
  - name          : "Hung Pham"
    affiliation   : " 4"
  - name          : "Christopher Sch\\\"{o}lzel"
    affiliation   : " 5"
  - name          : "S.H. Annabel Chen"
    affiliation   : " 1, 6, 7"
affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "Behavioural Science Institute, Radboud University, Netherlands"
  - id            : "3"
    institution   : "Departement de psychologie, Universite de Montreal, Canada"
  - id            : "4"
    institution   : "Eureka Robotics, Singapore"
  - id            : "5"
    institution   : "Life Science Informatics, THM University of Applied Sciences, Germany"
  - id            : "6"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "7"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
authornote: |
  * Correspondence concerning this article should be addressed to Dominique Makowski (HSS 04-18, 48 Nanyang Avenue, Singapore; dmakowski@ntu.edu.sg).
abstract: |
   NeuroKit2 is an open-source, community-driven, and user-friendly Python package dedicated to neurophysiological signal processing (e.g., ECG, EDA, EMG, ...). Its design philosophy is centred on user-experience and accessibility to both novice and advanced users. The package provides high-level functions that enable data processing in a few lines of code using validated pipelines, as well as low-level tools dedicated to specific processing steps. NeuroKit2 aims at improving transparency and reproducibility in neurophysiological research, as well as enabling exploration and innovation.  
keywords          : "Neurophysiology, Biosignals, Python, ECG, EDA, EMG"
wordcount         : ""
bibliography      : ["bibliography.bib"]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output:
  papaja::apa6_pdf:
    keep_tex: FALSE
    latex_engine: xelatex
  papaja::apa6_word:
    keep_tex: FALSE
header-includes:
   - \usepackage[labelfont=bf, font={color=gray,small}]{caption}
   - \usepackage{float}
   - \usepackage[document]{ragged2e}
editor_options:
  chunk_output_type: console
---

\justify

```{r r_setup, include = FALSE, warning=FALSE, message=FALSE}
library("papaja")
library("kableExtra")
options(knitr.kable.NA = 'None')

library(tidyverse)
library(easystats)

# Setup python - you need to change the path to your python distribution
library(reticulate)
reticulate::use_python("D:/Downloads/WPy64-3810/python-3.8.1.amd64/")
matplotlib <- import("matplotlib")
matplotlib$use("Agg", force = TRUE)
```



<!-- Research gap -->
Neurophysiological measurements increasingly gain popularity in the study of cognition and behavior. These measurements include electroencephalography (EEG), electrocardiography (ECG), electromyography (EMG) and electrodermal activity (EDA). Their popularity is driven by theoretical motivations [e.g., the growth of embodied or affective neuroscience; @Kiverstein2015] as well as practical reasons. The latter include low costs (especially compared with other imaging techniques, such as MRI or MEG), ease of use (e.g., portability, setup speed), and the increasing availability of recording devices [e.g., wearables; @yuehong2016internet]. Moreover, the extraction of meaningful information from neurophysiological signals is facilitated by current advances in signal processing algorithms [@clifton2012machine; @roy2019deep]. 

Unfortunately, these algorithms are mostly inaccessible to researchers without experience in programming and signal processing. Moreover, many software tools for neurophysiological analyses are limited to one type of signal (for instance, focused on ECG). This makes using those tools inconvenient for researchers who whish to analyse datasets containing multiple physiological modalities.

<!-- Response to gap -->
*NeuroKit2* adresses these challenges by offering a free, user-friendly, and comprehensive solution for neurophysiological data processing. It is an open-source Python package, developed by a multi-disciplinary team that actively invites new colaborators. *NeuroKit2* aims at being accessible, well-documented, well-tested, cutting-edge, flexible and efficient.

<!-- Accessibility and documentation -->
The package is implemented in Python 3 [@python3], which means that *NeuroKit2's* users benefit from an large amount of learning resources and a vibrant community. *NeuroKit2* depends on relatively few, well established and robust packages from the Python data analysis ecosystem [@scipy, @NumPy, @pandas, @SciPy, @scikit-learn, @MatplotLib].


*NeuroKit2's* source code is available under the permissive MIT license on GitHub (*https://github.com/neuropsychology/NeuroKit*). Its documentation (*https://neurokit2.readthedocs.io/*) is automatically built and rendered from the code and includes guides for installation and contribution, a description of the package's functions, as well as several "hands-on" examples and tutorials (e.g., how to extract and visualize individual heartbeats, how to analyze event-related data etc.). 


Users can add new examples by simply uploading a Python notebook [@kluyver2016jupyter] to the Github repository. The notebook will automatically be displayed on the website, ensuring easily accessible and evolving documentation. Moreover, users can try out the example notebooks directly in their browser via a cloud-based *Binder* environment [@Jupyter2018]. Finally, the issue tracker on GitHub offers a convenient and public forum that allows newcomers and potential collaborators to get help and insight into the development of the package.

<!-- Reliability and Evolution -->
*NeuroKit2* aims at being reliable and trustworthy, including peer-reviewed processing pipelines and functions tested against implementations of established software such as *BioSPPy* [@biosppy], *hrv* [*under review*](https://github.com/openjournals/joss-reviews/issues/1867), *PySiology* [@PySiology], *HeartPy* [@HeartPy], *systole* [@Systole] or *nolds* [@nolds]. The repository leverages a comprehensive test suite and continuous integration [e.g., travis CI] to ensure stability and prevent errors. Thanks to its collaborative and open development as well as its modular organization, *NeuroKit2* can continuously evolve, adapt, and integrate cutting-edge methods as they are emerging.


<!-- Powerful and flexible: API -->

# Design Philosophy
We strive for a flexible user interface (API) which offers both, high-level abstractions as well as fine-grained control. 
This allows us to adress the needs of beginners as well as experienced users. Beginning users can implement complex analyses in very few lines of code, while more experienced users can leverage low-level functionality. Concretely, our functions are implemented in three layers of abstraction.


## Low-level: Base Utilities for Signal Processing

The basic building blocks are functions for general signal processing, i.e., filtering, resampling, interpolation, peak detection, etc.. These functions are signal-agnostic, and include a lot of parameters. For instance, one can change the filtering -method, -frequencies, and -order. Most of these functions are based on established algorithms implemented in *scipy* [@scipy]. Examples of such functions include `signal_filter()`, `signal_interpolate()`, `signal_resample()`, `signal_detrend()`, and `signal_findpeaks()`.


## Mid-level: Neurophysiological Processing Steps

The base utilities are used by mid-level functions specific to the different physiological modalities (i.e., ECG, RSP, EDA, EMG, PPG). These functions carry out modality-specific signal processing steps, such as cleaning, peak detection, phase classification or rate computation. Critically, for each type of signal, the same function names are called (in the form `signaltype_functiongoal()`) to achieve equivalent goals, e.g., `*_clean()`, `*_findpeaks()`, `*_process()`, `*_plot()`, making the implementation intuitive and consistent across different modalities.

For example, the `rsp_clean()` function uses `signal_filter()` and `signal_detrend()`, with different sets of default parameters that can be switched with a "method" argument (corresponding to different published or validated pipelines). For instance, setting `method="khodadad2018"` will use the cleaning workflow described in @khodadad2018optimized. However, if a user wants to build their own custom cleaning pipeline, they can use the cleaning function as a template, and directly tweak the parameters in the low-level signal processing operations.


## High-level Wrappers for Processing and Analysis

The mid-level functions are assembled in high-level "master" functions, that are convenient entry points for new users. For instance, the `ecg_process()` function internally chains the mid-level functions `ecg_clean()`, `ecg_findpeaks()`, `ecg_rate()`. A specific processing pipeline can be selected with the `method` argument, that is then propagated throughout the internal functions. Easily switching between processing pipelines allows for the comparison of different methods, and streamlines critical but time-consuming steps in reproducible research, such as the validation of data preparation and quality control [@Quintana2016]. Finally, the package includes convenience meta-functions (e.g., `bio_process`) that enable the combined processing of multiple types of signals at once (e.g., `bio_process(ecg=ecg_signal, eda=eda_signal)`).

Performing a set of operations with sensible default parameters can be rewarding, especially for beginners, allowing them to perform cutting-edge processing or replication of research steps without requiring much programming expertise. Moreover, it contributes to the demystification of the usage of APIs (as opposed to GUI-based software such as *SPSS*, *Kubios*, or *Acqknowledge*), providing a welcoming framework to further explore physiological data processing at all levels of detail. Importantly, more advanced users can easily build custom analysis pipelines by using the mid-level functions, allowing for a finer control over the processing parameters. We believe that this implementation is a well-calibrated trade-off between flexibility and user-friendliness. 


# Examples

In this section, we present two examples that illustrate the most common use-cases. The first example is an event-related paradigm, in which the interest lies in short-term physiological changes related to specific events. The second example shows how to extract characteristic features of physiological activity during a longer period of time (not necessarily tied to a specific and sudden event). The example datasets are made available with the package and can be downloaded using the `data()` function.


## Event-related Paradigm

This example dataset contains ECG, RSP and EDA signals of one participant who was presented with four emotional images [from the NAPS database; @marchewka2014nencki], in a typical (albeit highly shortened) experimental psychology paradigm.

Signals are 2.5 minutes long and are recorded at a frequency of 100Hz [note that the sampling rate is low for storage purposes and should be higher in actual recordings, see @Quintana2016]. It has 4 channels including three physiological signals, and one corresponding to the marking of events with a photosensor (which signal decreases when a stimulus appeared on the screen).



```{python eventrelated_data, include=TRUE, results="hide", eval=TRUE, echo = TRUE}
# Load the package
import neurokit2 as nk

# Download the example dataset
data = nk.data("bio_eventrelated_100hz")

# Figure 1. Visualize 10 seconds of data (on the same scale)
nk.signal_plot(data[900:1900], standardize=True)

# Process the data
df, info = nk.bio_process(ecg=data["ECG"],
                          rsp=data["RSP"],
                          eda=data["EDA"],
                          sampling_rate=100)

# Find events
conditions = ["Negative", "Neutral", "Neutral", "Negative"]
events = nk.events_find(event_channel=data["Photosensor"],
                        threshold_keep='below',
                        event_conditions=conditions)

# Epoch the data
epochs = nk.epochs_create(data=df,
                          events=events,
                          sampling_rate=100,
                          epochs_start=-0.1,
                          epochs_end=4)

# Extract event related features
results = nk.bio_analyze(epochs)

# Show subset of results
results[["Condition", "ECG_Rate_Mean", "RSP_Rate_Mean", "EDA_Peak_Amplitude"]]
```

```{r include=TRUE, eval=TRUE, echo = FALSE, fig.width=10, fig.height=6, fig.cap="Subset of the dataset showing one event (in orange) and the other physiological signals."}
py$data %>%
  standardize() %>%
  mutate(Time = 1:n() / 100) %>%
  slice(900:1900) %>%
  pivot_longer(1:4) %>%
  mutate(name = fct_relevel(name, c("ECG", "RSP", "EDA", "Photosensor"))) %>%
  ggplot(aes(x=Time, y=value, color=name, size=name)) +
  geom_line() +
  theme_modern() +
  scale_color_manual(values=c("ECG"="red", "EDA"="#9C27B0", "RSP"="#2196F3", "Photosensor"="#FF9800")) +
  scale_size_manual(values=c("ECG"=0.66, "EDA"=2, "RSP"=2, "Photosensor"=2)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_blank(),
        legend.position = "top") +
  ylab("Time (s)") +
  guides(size = guide_legend(override.aes = list(size = 2)))
```
```{r table1word, eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# For word
knitr::kable(py$results[c("Condition", "ECG_Rate_Mean", "RSP_Rate_Mean", "EDA_Peak_Amplitude")], format="markdown", digits = 3, caption = "Subset of the ouput related to event-related analysis characterizing the pattern of physiological changes related to specific stimuli.", row.names = FALSE)
```
```{r table1pdf, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE, out.width = "\\textwidth", fig.pos = "!ht", results = "asis"}
# For PDFs
kable(py$results[c("Condition", "ECG_Rate_Mean", "RSP_Rate_Mean", "EDA_Peak_Amplitude")], format="latex", digits = 2, booktabs = TRUE, caption = "Subset of the ouput related to event-related analysis characterizing the pattern of physiological changes related to specific stimuli.", linesep="") %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```


In this example, after loading the package and the example dataset, each physiological signal is processed using `bio_process()`. Stimulus onsets in the photosensor are detected separately with `events_find()`. Once we have the preprocessed signals and the location of events, we can slice the data into segments corresponding to a time window (ranging from -0.1 to 4 seconds) around each stimulus with `epochs_create()`. Finally, relevant features are computed for each epoch (i.e., each stimulus) by passing them to `bio_analyze()`.

The features include for example the changes in rate of ECG and RSP signals (e.g. maximum, minimum and mean rate after stimulus onset, and the time at which they occur), and the peak characteristics of the EDA signal (e.g., occurrence of skin conductance response (SCR), and if SCR is present, its corresponding peak amplitude, time of peak, rise and recovery time). In addition, respiration and cardiac cycle phases are extracted (i.e., the respiration phase - inspiration/expiration - and cardiac phase - systole/diastole - occurring at the onset of event).

This example shows the straightforward process of extracting features of physiological responses. This pipeline can easily scale up to group-level analysis by aggregating the average of features across subjects. In addition to streamlining data analyses, *NeuroKit2* aims to provide researchers an extensive suite of signal features, allowing for precise interpretations in terms of relationship between physiological activity and neurocognitive processes. In this example (see **Table 1**), exposure to negative stimuli, as compared to neutral stimuli, is related to stronger cardiac deceleration, higher skin conductance response, and accelerated breathing rate (note that statistically, this illustrative interpretation is purely descriptive).



## Resting-state Features

The second dataset corresponds to 5 minutes of physiological activity of a human participant at rest (eyes-closed in a seated position), under no specific set of instructions. It contains three channels (ECG, PPG and RSP) sampled at a frequency of 100Hz.

```{python intervalrelated, include=TRUE, results="hide", eval=TRUE, echo = TRUE}
# Load the package
import neurokit2 as nk

# Download the example dataset
data = nk.data("bio_resting_5min_100hz")

# Process the data
df, info = nk.bio_process(ecg=data["ECG"],
                          rsp=data["RSP"],
                          sampling_rate=100)

# Extract features
results = nk.bio_analyze(df)

# Show subset of results
results[["ECG_Rate_Mean", "ECG_HRV_RMSSD", "RSP_Rate_Mean", "RSA_P2T_Mean"]]
```

```{r table2word, eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# For word
knitr::kable(py$results[c("ECG_Rate_Mean", "ECG_HRV_RMSSD", "RSP_Rate_Mean", "RSA_P2T_Mean")], format="markdown", digits = 3, caption = "Subset of properties characterizing the physiological activity over a period of 5 minutes of resting-state", row.names = FALSE)
```
```{r table2pdf, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE, out.width = "\\textwidth", fig.pos = "!ht", results = "asis"}
# For PDFs
kable(py$results[c("ECG_Rate_Mean", "ECG_HRV_RMSSD", "RSP_Rate_Mean", "RSA_P2T_Mean")], format="latex", digits = 2, booktabs = TRUE, caption = "Subset of properties characterizing the physiological activity over a period of 5 minutes of resting-state.", linesep="") %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```



In this example, the steps of the analysis are identical to the previous example, including loading the package, the dataset and processing the data. The difference is that there is no epoching, as we want to compute features related to the whole dataset. Thus, we can directly pass the dataframe to `bio_analyze()`, which will detect that these are not epochs, and compute the appropriate features accordingly. These include for instance the average heart and breathing rate, as well as indices of heart rate variability (HRV) and respiratory sinus arrhythmia (RSA).


This example illustrates a second type of physiological analysis, that we refer to as interval-related. Interval-related analyses compute features of signal variability and activation patterns over a longer-term period of time (typically minutes). *NeuroKit2* allows for the fast creation of a standardized and reproducible pipeline to describe this kind of physiological activity.



# Discussion

*NeuroKit2* offers neurophysiological signal processing software to people with all levels of programming experience. Its development is focused on an intuitive user-experience and a collaborative community. It is also a pragmatic answer to the broader need for transparent and reproducible methods in neurophysiology. Its modular organization not only facilitates the use of existing and validated processing pipelines, but is also suitable for experimentation and innovation.

We expect the packages' future evolution to be driven by the communities' needs and the advances in related fields. Possible directions include extending the support for other types of bodily signals (e.g., electrogastrography - EGG, electrooculography - EOG) and achieving performance gains for large datasets by using efficient algorithms. Further validation of the available processing pipelines could be made through the (re)analysis of public databases. In line with this objective, the support of standardized data structure formats (e.g. WFDB, BIDS, ...) could be extended.

In conclusion, we believe that *NeuroKit2* provides useful tools for anyone who is interested in analyzing data from wearables, "smart health devices", or research-grade hardware. This makes the package useful beyond academic research in neuroscience and psychology, including personal health monitoring and exercise science. By providing free, open-source solutions we hope to increase the autonomy of researchers and shorten the delay between data acquisition and analysis. Finally, we hope that *NeuroKit2* encourages users to become part of a supportive open-science community with diverse areas of expertise - rather than relying on closed-source and proprietary software. 


# Conflict of Interest

The authors declare that the research was conducted in the absence of commercial or financial relationships that could constitute a conflict of interest.

# Acknowledgements

We would like to thank Prof. C. F. Xavier for inspiration, all the contributors (https://neurokit2.readthedocs.io/en/latest/authors.html), and the users for their support.



\newpage

# References
```{r create_r-references}
r_refs(file = "bibliography.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
