---
title: '**NeuroKit2: A Python Toolbox for Neurophysiological Signal Processing**'
author: 
  - name          : "Dominique Makowski"
    affiliation   : " 1,*"
    corresponding : no    # Define only one corresponding author
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
  - name          : "Tam Pham"
    affiliation   : " 1"
  - name          : "Zen J. Lau"
    affiliation   : " 1"
  - name          : "Jan C. Brammer"
    affiliation   : " 2"
  - name          : "Hung Pham"
    affiliation   : " 3"
  - name          : "Francois Lesspinasse"
    affiliation   : " 4"
  - name          : "Christopher Sch\"{o}lzel"
    affiliation   : " 5"
  - name          : "S.H. Annabel Chen"
    affiliation   : " 1, 6, 7"
affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "???"
  - id            : "3"
    institution   : "???"
  - id            : "4"
    institution   : "???"
  - id            : "5"
    institution   : "???"
  - id            : "6"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "7"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
authornote: |
  * Correspondence concerning this article should be addressed to Dominique Makowski (HSS 04-18, 48 Nanyang Avenue, Singapore; dmakowski@ntu.edu.sg).
abstract: |
   The NeuroKit2 toolbox is an open-source Python package aimed at providing users with comprehensive and flexible functionality in neurophysiological signal processing. It developed from a collaborative project aimed at offering programming ease for both novice and advanced users to perform elaborate analyses of electrocardiogram (ECG), respiratory (RSP), electrodermal activity (EDA), and electromyography (EMG) data. It comprises of a consistent set of user-friendly, high-level functions that implements an all-in-one cleaning, preprocessing, and processing pipeline with sensible defaults. At the same time, greater flexibility and parametric control can be achieved by using Neurokit2's mid-level functions to build a custom analysis pipeline. (talk about novelty?)
keywords          : "neurophysiology, ECG, EDA, EMG, RSP"
wordcount         : ""
bibliography      : ["bibliography.bib"]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output: 
  papaja::apa6_pdf:
    keep_tex: FALSE
  papaja::apa6_word:
    keep_tex: FALSE
header-includes:
   - \usepackage[labelfont=bf, font={color=gray,small}]{caption}
   - \usepackage{float}
   - \usepackage[document]{ragged2e}
---

\justify

```{r setup, include = FALSE, warning=FALSE, message=FALSE}
library("papaja")
library("kableExtra")
```






<!-- Research gap -->
The field of cognitive neuroscience and psychology is increasingly relying on neurophysiological methods. One of the reasons is that such approaches often offer low monetary cost (especially compared with other imaging techniques, such as MRI) and high user convenience (e.g., portability). At the same time, the fields of signal processing and computational data science are growing strongly, tackling issues and limitations, and pushing the horizon of possibilities and opportunities. However, as these methods are often not easily accessible and user-friendly, neurophysiological data processing remains a challenge for many researchers without a formal programming training.

<!-- Response to gap -->
*NeuroKit2* aims at addressing this gap by offering a free and user-friendly solution for neurophysiological data processing. It is an open-source Python package, developed in a collaborative environment that continues to welcome contributors from different countries and fields. Historically, *NeuroKit2* is the re-forged successor *NeuroKit* (*https://github.com/neuropsychology/NeuroKit.py*), a PhD side project that ended up attracting a lot of users and success (236 GitHub stars as of 13-03-2020). The new version takes on its best features and design choices, and re-implements them in a professional and well-thaught way. It aims at being 1) accessible, 2) well-documented, 2) reliable, 4) cutting-edge and 5) powerful.

<!-- Accessibility and documentation -->
The package is available for Python [@python3] and thus benefits of its important base of users, existing tutorials and large online community. It is also relatively lightweight, enabling its use as a dependency in other software. The package source code is available under a permissive license on GitHub (*https://github.com/neuropsychology/NeuroKit*); along with its documentation, it is automatically built and hosted at *https://neurokit2.readthedocs.io/*. Apart from instructions on installation and contribution, and a decription of the package's functions, it also includes several "hands-on" examples and tutorials providing a walk-through on how to address specific issues (for instance, how to extract and visualize individual heartbeats). New examples can be easily added by users simply by uploading a Python notebook file to the repository. This notebook file will be automatically transformed into a webpage and displayed on the website, ensuring a state of the art and evolutive documentation. The accessibility for newcomers is reinforced by the issue tracker of GitHub, where users can create public issues to inquire for help. 

<!-- Reliability and Evolution -->
The packages is made to be reliable, and its functions are tested against existing implementations of established reference software such as *BioSPPy* [@biosppy], *hrv* [*under review*](https://github.com/openjournals/joss-reviews/issues/1867), *PySiology* [@PySiology], *HeartPy* [@HeartPy], *systole* [@Systole] or *nolds* [@nolds]. The code itself includes a comprehensive test suite to ensure stability and prevent error. Moreover, the issue tracker allows users to easily report any bugs and track their fixation. Thanks to its collaborative and open developpment, as well as its modular organization, *NeuroKit2* is being developped with a longterm perspective in mind, aiming at remaining cutting-edge through its ability to evolve, adapt, and integrate new methods as they are being developped.

<!-- - Powerful and flexible: API -->
Finally, we believe that the design philosophy contributes to a powerful (allowing to achieve a lot with very few functions) yet flexible (enabling fine control and precision over what is done) user interface. It is described below.

# Design Philosophy





## Organization

<!-- Neurokit2 was developed using Python drawing on well-established existing libraries such as **Numpy**, **Pandas**, **SciPy**, **MatplotLib** etc. As stated above, one of the primary goals of the package is to provide easy-to-use functions, catering to programming novices, but at the same time allowing maximal flexibility for advanced users. -->

3 levels of functions.


<!--  and it is well-documented with tutorials and examples. More importantly, the toolbox was designed to provide a consistent yet powerful and flexible API. It offers clear code organization across different signals (e.g., electrocardiogram (ECG), respiration (RSP), electrodermal activity (EDA), electromyography (EMG)) for different processing purposes. On one hand, it comprises of high-level functions that streamlines and incorporates all stages of signal processing, and on the other hand, mid-level functions (e.g., for cleaning data, extracting signal rate etc.) can be modified easily for user preference. Advanced users will find this useful for implementing their own algorithms and building their own custom analysis pipeline, giving them more parametric control. Hence, Neurokit2 is made accessible and reliable for users of various programming levels. -->

<!-- In this paper, we will provide an overview of the Neurokit2 package, its code organization and its various features. -->



<!-- One of the strengths of this package is its consistency. For each type of signals (ECG, RSP, EDA, EMG...), the same function names, in the form **signaltype_functiongoal()**, can be called to achieve equivalent goal. Therefore, in each signal-specific module, users can expect to find functions with similar names such as **signaltype_clean()**, **signaltype_findpeaks()**, **signaltype_plot()** or **signaltype_process()**. Following, we will elaborate on the functions that are signal-specific but designed to perform the same processing task in the pipeline. -->

## Processing and analysis



This design withholds the promise of being able to do a full processing and features extraction of your data with only 2 functions.





<!-- The package comprises of multiple modules: ecg, rsp, eda, emg, ppg, signal, eeg, epochs, events, complexity, bio, misc and stats. -->

<!-- The ecg, rsp, eda, and emg modules are tailored to the processing and analysis of the respective signals. They encompass mid-level functions and each performs a specific part of the processing pipeline such as *cleaning* to increase the signal-to-noise ratio, *extraction of information* to better quantify the signal, *plotting* to visualize the extracted features on the signal, and *extraction of features* for event-related analysis on epochs of signal or period-related analysis for signal with longer duration. Generally, users can start writing their processing scripts by calling the mid-level functions one-by-one, according to the order of steps that they have in their processing pipelines. And if users do not have a specific pipeline in mind, NeuroKit2 provides a single high-level function in each module called **_process** to perform every step in one command line, following the Neurokit2 default pipeline. -->

<!-- Besides the signal-specific modules, there are modules with functions that also perform specific steps in the processing pipeline but are non-specific to any type of signal. Such as *events* to support incorporation and visualization of event markers in the signal or *epochs* to chunk signals into event-related segments. -->

<!-- Usually not as user-friendly as the aforementioned modules, there are modules with low-level functions, such as signal, complexity, misc and stats, that support the implementation of core functionality in other modules. Unless the users wish to, they do not need to call these low-level functions into the processing scripts. -->

<!-- Last but not least, there is a high-level function in the bio module that allows processing of all types of signals at once. As powerful as it might sound, this function basically combines the high-level function of each signal type into one script, allowing users who do not wish to stray from Neurokit2 default processing pipeline to perform all steps in one command line. -->

<!-- ```{r table, echo=FALSE} -->
<!-- ecg = c("ecg_clean()", "ecg_delineate()", "ecg_eventrelated()", "ecg_findpeaks()", "ecg_findpeaks()", "ecg_fixpeaks()", "ecg_peaks()", "ecg_plot()", "ecg_process()", "ecg_rate()", "ecg_rate()", "ecg_rsp()", "ecg_simulate()") -->
<!-- events = c("events_find()", "events_plot()", "events_to_mne()") -->
<!-- # write.csv(table, "figures/table.csv", row.names = FALSE) -->
<!-- ``` -->



# Example


Despite not having a Graphical User Interface (GUI), NeuroKit2 is accessible to people with very little knowledge of python or programming in general. 


## Event-related Paradigm

```{python, eval=FALSE, echo = TRUE}
# Imports
import neurokit2 as nk

# Download example dataset
data = pd.read_csv("https://raw.githubusercontent.com/neuropsychology/NeuroKit/master/data/example_bio_100hz.csv")

# Process the data
df, info = nk.bio_process(ecg=data["ECG"], rsp=data["RSP"], eda=data["EDA"], emg=None, sampling_rate=100)

# Find events
events = nk.events_find(event_channel=data["Photosensor"], threshold_keep='below',
event_conditions=["Negative", "Neutral", "Neutral", "Negative"])

# Visualize events in the signal
plot = nk.events_plot(events=events, signal=data)

# Epoch the data
epochs = nk.epochs_create(data=df, events=events, sampling_rate=100, epochs_start=-0.1, epochs_end=1.9)

# Extract event related fetures
ecg_event = nk.ecg_eventrelated(epochs)
rsp_event = nk.rsp_eventrelated(epochs)
eda_event = nk.eda_eventrelated(epochs)
```

**ADD HERE THE EXAMPLE ABOUT EVENT RELATED FEATURES using the example dataset**


## Resting-state Features 


**ADD here an example on resting feature dataset (to select from our data, downsample, transform to csv and add to the datasets)**







# Conflict of Interest

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

# Acknowledgements

All the contributors (https://neurokit2.readthedocs.io/credits.html) that reported bugs, and the users.








\newpage

# References
```{r create_r-references}
r_refs(file = "bibliography.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
